---
title: "Get Started"
author: "Kieran Healy"
date: "10-January-2020"
output: html_document
---

## Load Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gapminder)
library(here)
library(tidyverse)
library(socviz)
```

A code chunk:

```{r 02-get-started-1}

```

## Things to know about R

### Everything has a name

```{r 02-get-started-2}

```


### Everything is an object

```{r objects}
c(1, 2, 3, 1, 3, 5, 25)

```

```{r create_objects}
my_numbers <- c(1, 2, 3, 1, 3, 5, 25)

your_numbers <- c(5, 31, 71, 1, 3, 21, 6)
```

```{r 02-get-started-3}
my_numbers
```


### You do things using functions

```{r 02-get-started-4}
my_numbers
```


```{r this_will_error, eval = FALSE}
mean()
```

```{r 02-get-started-5}
mean(x = my_numbers)

mean(x = your_numbers)
```

```{r 02-get-started-6}
mean(my_numbers)
```

```{r 02-get-started-7}
my_summary <- summary(my_numbers)
```


```{r 02-get-started-8}
my_summary
```


### Functions come in libraries

```{r output}

table(my_numbers)

sd(my_numbers)

my_numbers * 5

my_numbers + 1

my_numbers + my_numbers

```

### If you're not sure what an object is, ask for its class

```{r getclass}
class(my_numbers)
class(my_summary)
class(summary)
```


```{r 02-get-started-9}
my_new_vector <- c(my_numbers, "Apple")
my_new_vector
class(my_new_vector)
```


```{r titanic1}
titanic
class(titanic)
```

```{r titanic2}
titanic$percent
```


```{r titanic3}
titanic_tb <- as_tibble(titanic)
titanic_tb
```


### To see inside an object, ask for its structure, or use RStudio's object inspector

```{r str}

str(my_numbers)
str(my_summary)

```


## Be patient with R, and with yourself

Here are three very specific things to watch out for:
  
- Make sure parentheses are balanced and that every opening "`(`" has
  a corresponding closing "`)`".
- Make sure you complete your expressions. If you think you have completed typing your code, but instead of seeing the `>` command prompt at the console you see the `+`
  character instead, that may mean R thinks you haven't written a complete
  expression yet. You can hit `Esc` or `Ctrl-C` to force your way back to the console and try typing your code again. 
- In ggplot specifically, as you will see, we will build up plots a
  piece at a time by adding expressions to one another. When doing
  this, make sure your `+` character goes at the end of the line, and
  not the beginning. That is, write this:


```{r 02-get-started-10, echo=TRUE, eval=FALSE}
ggplot(data = mpg, aes(x = displ, y = hwy)) +
    geom_point()
```


and not this:

```{r 02-get-started-11, echo=TRUE, eval=FALSE}
ggplot(data = mpg, aes(x = displ, y = hwy))
+ geom_point()
```


## Get data into R

Remotely: 

```{r 02-get-started-12, echo = TRUE, eval = FALSE}

url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"

organs <- read_csv(file = url)
```


Or locally:

```{r 02-get-started-13}

organs <- read_csv(file = "data/organdonation.csv")


organs
```

## Make your first figure 

```{r gapminder-example-1}
gapminder
```


```{r first_plot, fig.height=6, fig.width=10, fig.cap="Life expectancy plotted against GDP per capita for a large number of country-years."}

p <- ggplot(data = gapminder,
            mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + 
  geom_smooth(mapping = aes(color = continent, fill = continent)) + 
  scale_x_log10(labels = scales::dollar)

```
