---
title: "Supplementary Material"
author: "Kieran Healy"
date: "10-January-2020"
output: html_document
---

## Load Libraries


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(gapminder)
library(here)
library(tidyverse)
library(ggrepel)
library(socviz)
```



## The basics of accessing and selecting things

```{r 09-supplementary-material-1}
my_numbers <- c(1, 2, 3, 1, 3, 5, 25)
your_numbers <- c(5, 31, 71, 1, 3, 21, 6)
```

```{r 09-supplementary-material-2}
my_numbers[4]
my_numbers[7]
```

```{r 09-supplementary-material-3}
my_numbers[2:4]
```

```{r 09-supplementary-material-4}
my_tb <- tibble(
    mine = c(1,4,5, 8:11),
    yours = c(3,20,16, 34:31))

class(my_tb)
my_tb
```

```{r 09-supplementary-material-5}
my_tb[3,1] # Row 3 Col 1
my_tb[1,2] # Row 1, Col 2 
```


```{r 09-supplementary-material-6}
my_tb[3,"mine"] # Row 3 Col 1
my_tb[1,"yours"] # Row 1, Col 2 
```

```{r 09-supplementary-material-7}
my_tb[3,"mine"] # Row 3 Col 1
my_tb[1,"yours"] # Row 1, Col 2 
```



```{r 09-supplementary-material-8}
my_tb[,"mine"] # All rows, Col 1
```


```{r 09-supplementary-material-9}
my_tb[4,] # Row 4, all cols
```


```{r 09-supplementary-material-10}
my_tb$mine
```


```{r 09-supplementary-material-11}
out <- lm(mine ~ yours, data = my_tb)

out$coefficients

out$call

out$qr$rank # nested 
```


```{r 09-supplementary-material-12}
my_tb$ours <- my_tb$mine + my_tb$yours
my_tb
```

### Tidy data

```{r 09-supplementary-material-13}
knitr::kable(preg, caption="Some untidy data.", booktabs = TRUE)
```

```{r 09-supplementary-material-14}
knitr::kable(preg2,
             caption="The same data, still untidy, but in a different way.", booktabs = TRUE)
```

```{r 09-supplementary-material-15}
preg3 <- preg %>% 
  gather(treatment, n, treatmenta:treatmentb) %>%
  mutate(treatment = gsub("treatment", "", treatment)) %>%
    arrange(name, treatment)
knitr::kable(preg3, 
      caption="Tidied data. Every variable a column, every observation a row.", booktabs = TRUE)
```

Recent changes to `tidyr` introduce `pivot_longer()` and `pivot_wider()` as friendlier alternatives to `gather()` and `spread()`.

```{r 09-supplementary-material-15-pivot}
preg4 <- preg %>% 
  pivot_longer(treatmenta:treatmentb, 
               names_to = "treatment", 
               names_prefix = "treatment",
               values_to = "n") %>%
    arrange(name, treatment)
knitr::kable(preg4, 
      caption="Tidied data using pivot_longer(). Every variable a column, every observation a row.", booktabs = TRUE)
```

Another example.


```{r 09-supplementary-material-16}
edu
```

First with `gather()`:

```{r 09-supplementary-material-18, echo=TRUE, message=FALSE, tidy = FALSE}
edu_t <- gather(data = edu,
                key = school,
                value = freq,
                elem4:coll4)

head(edu_t) 
tail(edu_t) 
```

And now equivalently, with `pivot_longer()`:

```{r 09-supplementary-material-18-pivot, echo=TRUE, message=FALSE, tidy = FALSE}
edu_t <- pivot_longer(data = edu, 
                      cols = elem4:coll4, 
                      names_to = "school",
                      values_to = "freq")
  
head(edu_t) 
tail(edu_t) 
```

(Note the ordering of the rows in the resulting tibble is slightly different from the one produced by `gather`, but the operations are otherwise equivalent.)

## Common problems reading in data

### Date formats

```{r 09-supplementary-material-19, fig.caption="A bad date.", fig.cap = "A bad date."}
p <- ggplot(data = bad_date, aes(x = date, y = N))
p + geom_line()
```


```{r 09-supplementary-material-20, echo = TRUE, fig.width = 5, fig.height = 4, fig.cap = "Still bad."}

bad_date2 <- rbind(bad_date, bad_date)

p <- ggplot(data = bad_date2, aes(x = date, y = N))
p + geom_line()

```


```{r 09-supplementary-material-21}
# install.packages("lubridate")
library(lubridate)

bad_date$date <- mdy(bad_date$date)
head(bad_date)

```

```{r 09-supplementary-material-22, fig.cap="Much better.", out.width="100%", fig.height=4, fig.width=6, fig.margin=TRUE}
p <- ggplot(data = bad_date, aes(x = date, y = N))
p + geom_line()
```


### Year-only dates

```{r 09-supplementary-material-23, eval = FALSE}

url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"

bad_year <- read_csv(url)
bad_year %>% select(1:3) %>% sample_n(10)

```

```{r 09-supplementary-material-24}

bad_year <- read_csv(file = "data/organdonation.csv")
bad_year %>% select(1:3) %>% sample_n(10)

```

```{r 09-supplementary-material-25, fig.cap="Integer year shown with a decimal point.",  fig.height = 4, fig.width = 5}

p <- ggplot(data = bad_year, aes(x = year, y = donors))
p + geom_point()

```


```{r 09-supplementary-material-26}
bad_year$year <- int_to_year(bad_year$year)
bad_year %>% select(1:3)
```

## Visualizing Missing Data

```{r missing-1}
drat::addRepo("kjhealy")
install.packages("congress")
library(congress)

library(naniar)
library(visdat)

```

```{r missing-2}
vis_dat(congress)
```

```{r missing-3}
gg_miss_var(organdata)
```

```{r missing-4}
vis_dat(organdata)
```

```{r missing-5}
miss_var_summary(organdata)
```

```{r missing-6}
organdata %>%
  select(consent_law, year, pubhealth, roads) %>%
  group_by(consent_law) %>%
  miss_var_summary()

```

```{r missing-7}

vis_miss(organdata)
```

```{r missing-8}

gg_miss_upset(organdata)
```

```{r missing-9}
ggplot(data = organdata, mapping = aes(x = pubhealth, y = donors)) + 
geom_point()
```

```{r missing-10}

ggplot(data = organdata, mapping = aes(x = pubhealth, y = donors)) + 
geom_miss_point()
```

```{r missing-11}
gg_miss_fct(x = riskfactors, fct = marital)
```

## Zero Counts in Dplyr

```{r zero-1}

## Hex colors for sex
sex_colors <- c("#E69F00", "#993300")

## Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A")

## Group labels
mf_labs <- tibble(M = "Men", F = "Women")

df <- read_csv("data/fc_sample.csv")

df

```

```{r zero-2}
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N))


```

```{r zero-3}

df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               fill = sex)) +
    geom_col() +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_manual(values = sex_colors, 
                      labels = c("Women", "Men")) +
    labs(x = "Year", y = "Percent", fill = "Group") +
    facet_wrap(~ party)
```


```{r zero-4}

df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               color = sex)) +
    geom_line(size = 1.1) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_manual(values = sex_colors, 
                       labels = c("Women", "Men")) +
    guides(color = guide_legend(reverse = TRUE)) +
    labs(x = "Year", y = "Percent", color = "Group") +
    facet_wrap(~ party)
```


```{r zero-5}
df_f <- df %>% modify_if(is.character, as.factor)

df_f %>%
    group_by(start_year, party, sex) %>%
    tally()

```

```{r zero-5a}
df_f %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ggplot(aes(x = start_year,
               y = freq,
               color = sex)) +
    geom_line(size = 1.1) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_manual(values = sex_colors, 
                       labels = c("Women", "Men")) +
    guides(color = guide_legend(reverse = TRUE)) +
    labs(x = "Year", y = "Percent", color = "Group") +
    facet_wrap(~ party)
```

```{r zero-6}
df %>%
    group_by(start_year, party, sex) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N)) %>%
    ungroup() %>%
    complete(start_year, party, sex,
             fill = list(N = 0, freq = 0))

```

```{r zero-7}


```

## Network Data

```{r net1}
library(ggraph)
library(tidygraph)
library(graphlayouts)
library(janitor)

drat::addRepo("kjhealy")
install.packages("kjhnet")
library(kjhnet)

head(revere)

```


```{r net2}
revere_groups
```


```{r net3}
revere_groups %>%
  ggraph(layout = "kk") +
  geom_edge_link(aes(width = value), color = "gray80") +
  geom_node_label(aes(label = name))

```


```{r net4}
revere_persons
```


```{r net5}
revere_persons %>%
  mutate(centrality = centrality_eigen()) %>%
  ggraph(layout = "stress") +
  geom_edge_link0(mapping = aes(filter = value > 1, 
                                 width = value), 
           color = "gray90") +
  geom_node_point() +
  geom_node_label(mapping = aes(filter = centrality > 0.9, 
                                label = name), 
                  size = rel(2.5))
```


```{r net6}
il_tidy
```

```{r net7}
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree()) %>%
  ggplot(mapping = aes(x = centrality)) + 
  geom_histogram() + 
  labs(x = "Count of Victories", y = "Count of Warriors ") +
  theme_minimal() + 
  theme(legend.position = "top")

```


```{r net8}
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree()) %>%
  arrange(desc(centrality)) %>%
  top_n(10, wt = centrality) %>%
  ggplot(mapping = aes(x = centrality, 
                       y = reorder(name, centrality), 
                       color = affil)) + 
  geom_point(size = 3) + 
  labs(x = "Centrality", y = NULL, color = "Side") +
  theme_minimal() + 
  theme(legend.position = "top")

```


```{r net9}
il_tidy %>%
  activate(nodes) %>%
  ggraph(layout = "fr") +
  geom_edge_link(color = "gray80") +
  geom_node_point(aes(color = affil)) +
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("Athenian", "Trojan")) +
  guides(color = guide_legend(title = "Side", )) +
  labs(title = "Killings in The Iliad") +
  theme(plot.title = element_text(size = rel(3)))

```


```{r net10}
label_colors <- c(prismatic::clr_lighten("blue", 0.7), 
                  prismatic::clr_lighten("red", 0.7))
il_tidy %>%
  activate(edges) %>%
  filter(act == "kills") %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree(mode = "out")) %>%
  ggraph(layout = 'graphopt') +
  geom_edge_link(aes(start_cap = label_rect(node1.name),
                     end_cap = label_rect(node2.name)),
                 arrow = arrow(length = unit(1.5, 'mm'))) +
  geom_node_point(aes(color = affil)) +
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("Athenian", "Trojan")) +
  guides(color = FALSE, fill = FALSE) +
  geom_node_label(aes(filter = centrality > 3, 
                      label = name, fill = affil), 
                      size = rel(2.5)) +
  scale_fill_manual(values = label_colors) + 
  labs(title = "Killings in The Iliad") +
  theme(plot.title = element_text(size = rel(3)))

```


```{r net11}
il_tidy %>%
  activate(nodes) %>%
  mutate(centrality = centrality_degree(),
         betweenness = centrality_betweenness()) %>%
  arrange(desc(betweenness)) %>%
  top_n(10, wt = betweenness) %>%
  ggplot(mapping = aes(x = betweenness, 
                       y = reorder(name, 
                       betweenness), color = affil)) + 
  geom_point(size = 3) + 
  labs(x = "Betweenness", y = NULL, color = "Side") +
  theme_minimal() + 
  theme(legend.position = "top")

```

```{r net12}
il_tidy %>%
  activate(edges) %>%
  filter(act == "kills") %>%
  reroute(from = to, to = from) %>%
  activate(nodes) %>%
  mutate(alpha = centrality_alpha()) %>%
  arrange(desc(alpha)) %>%
    top_n(10, wt = alpha) %>%
  ggplot(mapping = aes(x = alpha, 
                       y = reorder(name, alpha), 
                       color = affil)) + 
  geom_point(size = 3) + 
  labs(x = "Alpha Centrality", y = NULL, color = "Side") +
  theme_minimal() + 
  theme(legend.position = "top")

```


## Functions

```{r addnum1, echo = FALSE}
add_xy <- function(x = NULL, y = NULL) {
    x + y
}

```

```{r addnum2}

add_xy(x = 1, y = 7)

```

```{r addnum4}
add_xy(x = 5, y = 2)
```


```{r plot-section-function-1, echo=TRUE, tidy = FALSE}

plot_section <- function(section="Culture", x = "Year",
                         y = "Members", data = asasec,
                         smooth=FALSE){
    require(ggplot2)
    require(splines)
    # Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            mapping = aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = "#999999",
                              size = 1.2, method = "lm",
                              formula = y ~ ns(x, 3)) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
    } else {
    p0 <- p + geom_line(color= "#E69F00", size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        labs(title = section)
    }

    print(p0)
}

```

```{r 09-supplementary-material-27, fig.cap='Using a function to plot your results.', out.width="50%", fig.width=4, fig.height=3, fig.show = "hold"}

plot_section("Rationality")
plot_section("Sexualities", smooth = TRUE)

```

```{r plot-section-function-2, echo = TRUE, tidy = FALSE}

plot_section <- function(section="Culture", x = "Year",
                         y = "Members", data = asasec,
                         smooth=FALSE, ...){
    require(ggplot2)
    require(splines)
    # Note use of aes_string() rather than aes() 
    p <- ggplot(subset(data, Sname==section),
            mapping = aes_string(x=x, y=y))

    if(smooth == TRUE) {
        p0 <- p + geom_smooth(color = "#999999",
                              size = 1.2, ...) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
    } else {
    p0 <- p + geom_line(color= "#E69F00", size=1.2) +
        scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
        labs(title = section)
    }

    print(p0)
}


```

```{r 09-supplementary-material-28, fig.cap='Our custom function can now pass arguments along to fit different smoothers to Section membership data.', out.width="50%", fig.width=4, fig.height=3, fig.show = "hold"}

plot_section("Comm/Urban",
             smooth = TRUE,
             method = "loess")
plot_section("Children",
             smooth = TRUE,
             method = "lm",
             formula = y ~ ns(x, 2))

```


## Preparing the county-level maps

```{r 09-supplementary-material-29, eval = FALSE, echo = TRUE}

# You will need to use install.packages() to install
# these map and GIS libraries if you do not already
# have them.

library(maptools)
library(mapproj)
library(rgeos)
library(rgdal)

us_counties <- readOGR(dsn="data/geojson/gz_2010_us_050_00_5m.json",
                       layer="OGRGeoJSON")

us_counties_aea <- spTransform(us_counties,
                    CRS("+proj=laea +lat_0=45 +lon_0=-100 \
                         +x_0=0 +y_0=0 +a=6370997 +b=6370997 \
                         +units=m +no_defs"))

us_counties_aea@data$id <- rownames(us_counties_aea@data)
```


```{r 09-supplementary-material-30, eval = FALSE, echo = TRUE}

alaska <- us_counties_aea[us_counties_aea$STATE == "02",]
alaska <- elide(alaska, rotate=-50)
alaska <- elide(alaska, scale=max(apply(bbox(alaska), 1, diff)) / 2.3)
alaska <- elide(alaska, shift=c(-2100000, -2500000))
proj4string(alaska) <- proj4string(us_counties_aea)

hawaii <- us_counties_aea[us_counties_aea$STATE=="15",]
hawaii <- elide(hawaii, rotate=-35)
hawaii <- elide(hawaii, shift=c(5400000, -1400000))
proj4string(hawaii) <- proj4string(us_counties_aea)

us_counties_aea <- us_counties_aea[!us_counties_aea$STATE %in% c("02", "15", "72"),]
us_counties_aea <- rbind(us_counties_aea, alaska, hawaii)

```

Finally, we tidy the spatial object into a data frame that ggplot can use, and clean up the `id` label by stripping out a prefix from the string. 

```{r 09-supplementary-material-31, eval = FALSE, echo = TRUE, tidy = FALSE}
county_map <- tidy(us_counties_aea, region = "GEO_ID")
county_map$id <- stringr::str_replace(county_map$id,
                                      pattern = "0500000US", replacement = "")
```



```{r 09-supplementary-material-32, eval = FALSE, echo = TRUE, tidy = FALSE}
county_map <- tidy(us_counties_aea, region = "GEO_ID")
county_map$id <- stringr::str_replace(county_map$id,
                                      pattern = "0500000US", replacement = "")
```


For more detail and code for the merge, see `github.com/kjhealy/us-county`.

## Plot themes and map themes

```{r 09-supplementary-material-33, eval = FALSE}
devtools::install_github("hrbrmstr/hrbrthemes")
```


```{r 09-supplementary-material-34, eval = FALSE, tidy = FALSE}

theme_map <- function(base_size=9, base_family="") {
    require(grid)
    theme_bw(base_size=base_size, base_family=base_family) %+replace%
        theme(axis.line=element_blank(),
              axis.text=element_blank(),
              axis.ticks=element_blank(),
              axis.title=element_blank(),
              panel.background=element_blank(),
              panel.border=element_blank(),
              panel.grid=element_blank(),
              panel.spacing=unit(0, "lines"),
              plot.background=element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0)
              )
}

```

